---
title: "Assignment 4 - Heart rate, respiration and interpersonal coordination"
author: "Louise, Anna, Oliver & Malte"
date: "November, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 4 - Heart rate, respiration and interpersonal coordination

Physiological data (here heart rate [variability], and respiration) are increasingly popular. Historically treated as pernicious noise to be regressed out of neuro-imaging data, there is now increasing research on how these signals tell us something important about cognition and beyond being just a signal of cognitive processes also impact them in interesting ways. Advanced sport science, and the quantified self movement (closely followed by marketing and communication) have hailed continuous physiological tracking as a powerful way to access and modify attitudes, habits, and performance. Further, as team coordination (in the military, in decision processes and organizational contexts) is more and more in focus, research has attempted to measure how interpersonal coordination between physiological systems might tell us something important about e.g. emotional and cognitive coordination. See references in the reading list for more on this.

In this assignment, you will learn to:
- collect physiological data
- pre-process physiological data (and grow further your mad R skills)
- model the continuous interdependence between two signals (using a multilevel model as proxy for a dynamical system approach)
- conservatively assess the presence of coordination between to signals in a controlled context

This assignment has two parts. The first part familiarizes you with heart rate, and respiration data and their preprocessing. The second part explores how to analyze interpersonal coordination of these signals.

These are the questions you need to be able to answer at the end of the assignment (aka that you need to submit as part of the portfolio)

1) How do you preprocess heart rate and respiration data? Describe the process. If any data needs to be excluded, list the excluded data and motivate the exclusion.

2) Do you observe interpersonal coordination in heart rate and respiration? Describe your control baseline, the method used to quantify coordination, and the statistical models used to infer whether coordination was higher than in the baseline. Report the results of the models.

3) Do you observe differences in coordination between conditions? Report the models and results.

4) Is respiration coordination a likely driver of heart rate coordination? Describe how you would test for it. Bonus points if you actually run the tests and report methods and results.

N.B. to give you a bit more data I included data from previous years (Study1, Study2 and Study 3). Note that synchronouns and turn-taking are the same across both studies, but the third condition is different: in the first year it was self-paced joint reading; in the second year it was the tv-series conversation.

## Let's get started

### Exploring physiological signals

- Choose one pair (one pair, three conditions)
- Load the logs
- Produce a plot of the participants' respiration signal and a different one of the participants' HR signal.
  N.B: remember the slides: artifacts, downsampling, scaling.
  N.B. The gridExtra::grid.arrange() function allows you to display the plots side by side. E.g. grid.arrange(plot1, plot2, plot3, ncol=3). There are also smarter packages, like cowplot and ggpubr.
- Can you eye-ball which condition if any displays more physiological coordination?

### First we read one data file and identify the procedure
- Load the file
- correctly identify all columns
- plot the data
- deal with the artifacts
- downsample the dat
- Add a column for study, group, trial and condition

```{r}
# Load the libraries
library(pacman, tidyverse)
p_load(tidyverse)

```

```{r}
# Load the file
G1T1 <- read_csv("data/Study1_G1_T1_Synchronous.csv")
G1T2 <- read_csv("data/Study1_G1_T2_TurnTaking.csv")
G1T3 <- read_csv("data/Study1_G1_T3_SelfPaced.csv")
```

```{r}
# Plot
p1HR <- ggplot(G1T1) +
  geom_line(aes(time, HR1, color = "red"))+
  geom_line(aes(time, HR2, color = "blue"))

p1RESP <- ggplot(G1T1) +
  geom_line(aes(time, Resp1, color = "red"))+
  geom_line(aes(time, Resp2, color = "blue"))

p2HR <- ggplot(G1T2) +
  geom_line(aes(time, HR1, color = "red"))+
  geom_line(aes(time, HR2, color = "blue"))

p2RESP <- ggplot(G1T2) +
  geom_line(aes(time, Resp1, color = "red"))+
  geom_line(aes(time, Resp2, color = "blue"))

p3HR <- ggplot(G1T3) +
  geom_line(aes(time, HR1, color = "red"))+
  geom_line(aes(time, HR2, color = "blue"))

p3RESP <- ggplot(G1T3) +
  geom_line(aes(time, Resp1, color = "red"))+
  geom_line(aes(time, Resp2, color = "blue"))

#(p1HR + p1RESP) / (p2HR + p2RESP) / (p3HR + p3RESP)

```

```{r}
## Remove outliers

#Defining function to get the previous non-outlier value in the time-series
getClosestMin <- function(i, withinThresh)
{
  if(i > max(withinThresh))
  {
    return(max(withinThresh))
  }
  for(j in 1:length(withinThresh))
  {
    if(withinThresh[j] > i){
      return(withinThresh[j - 1])
    }
  }
}

getClosestMax <- function(i, withinThresh)
{
  if(i < min(withinThresh))
  {
    return(min(withinThresh))  
  }
  
  for(j in length(withinThresh):1)
  {
    if(withinThresh[j] < i)
      return(withinThresh[j + 1])
  }
}
## Remove outliers
removeOuts <- function(ts){
  OutVals <- boxplot(ts)$out #Taking all outlier values according to the IQR
  withinThresh <- which(!(ts %in% OutVals)) #Taking all the indices that are NOT outliers
  for (i in 1:length(ts)){
    if(i %% 10000 == 0){ #Progress print
      print(i)
    }
    if(!(i %in% withinThresh))
    {
      if(i < withinThresh[1]){ 
        ts[i] = ts[getClosestMax(i, withinThresh)] #If the first value is an outlier find the next non-outlier value
      }
      else if(i == length(ts)){ 
        ts[i] = ts[getClosestMin(i, withinThresh)] #If the last value is an outlier find the previous non-outlier value
      }
      else{
        ts[i] = (ts[getClosestMin(i, withinThresh)] + ts[getClosestMax(i, withinThresh)]) / 2 #If the value is an outlier take the previous and the next non-outlier value and take the mean of these
      }
    }
  }
  return(ts)
}
```

```{r}
#Loading/Creating the proper variables
T1filename <- "T1noOut.csv"
T2filename <- "T2noOut.csv"
T3filename <- "T3noOut.csv"

if (file.exists(c(T1filename, T2filename, T3filename))){
  T1noOut <- read_csv(T1filename)
  T2noOut <- read_csv(T2filename)
  T3noOut <- read_csv(T3filename)
  print(paste("Loaded files:", T1filename, T2filename, T3filename))
}else{
    T1noOut <- G1T1 %>% mutate_at(c("HR1", "HR2", "Resp1", "Resp2"), removeOuts)
    T2noOut <- G1T2 %>% mutate_at(c("HR1", "HR2", "Resp1", "Resp2"), removeOuts)
    T3noOut <- G1T3 %>% mutate_at(c("HR1", "HR2", "Resp1", "Resp2"), removeOuts)
    write_csv(T1noOut, T1filename)
    write_csv(T2noOut, T2filename)
    write_csv(T3noOut, T3filename)
}


```


```{r}
# Plot raw data againt those with the artifacts removed
pOutvsNout1 <- ggplot()+
  geom_line(data = G1T1, aes(time, HR1, color = "red"))+
  geom_line(data = T1noOut, aes(time, HR1, color = "blue"))

pOutvsNout2 <- ggplot()+
  geom_line(data = G1T1, aes(time, HR2, color = "red"))+
  geom_line(data = T1noOut, aes(time, HR2, color = "blue"))

pOutvsNout3 <- ggplot()+
  geom_line(data = G1T1, aes(time, Resp1, color = "red"))+
  geom_line(data = T1noOut, aes(time, Resp1, color = "blue"))

pOutvsNout4 <- ggplot()+
  geom_line(data = G1T1, aes(time, Resp2, color = "red"))+
  geom_line(data = T1noOut, aes(time, Resp2, color = "blue"))

pOutvsNout5 <- ggplot()+
  geom_line(data = G1T2, aes(time, HR1, color = "red"))+
  geom_line(data = T2noOut, aes(time, HR1, color = "blue"))

pOutvsNout6 <- ggplot()+
  geom_line(data = G1T2, aes(time, HR2, color = "red"))+
  geom_line(data = T2noOut, aes(time, HR2, color = "blue"))

pOutvsNout7 <- ggplot()+
  geom_line(data = G1T2, aes(time, Resp1, color = "red"))+
  geom_line(data = T2noOut, aes(time, Resp1, color = "blue"))

pOutvsNout8 <- ggplot()+
  geom_line(data = G1T2, aes(time, Resp2, color = "red"))+
  geom_line(data = T2noOut, aes(time, Resp2, color = "blue"))

pOutvsNout9 <- ggplot()+
  geom_line(data = G1T3, aes(time, HR1, color = "red"))+
  geom_line(data = T3noOut, aes(time, HR1, color = "blue"))

pOutvsNout10 <- ggplot()+
  geom_line(data = G1T3, aes(time, HR2, color = "red"))+
  geom_line(data = T3noOut, aes(time, HR2, color = "blue"))

pOutvsNout11 <- ggplot()+
  geom_line(data = G1T3, aes(time, Resp1, color = "red"))+
  geom_line(data = T3noOut, aes(time, Resp1, color = "blue"))

pOutvsNout12 <- ggplot()+
  geom_line(data = G1T3, aes(time, Resp2, color = "red"))+
  geom_line(data = T3noOut, aes(time, Resp2, color = "blue"))

pOutvsNout1
pOutvsNout2
pOutvsNout3
pOutvsNout4
pOutvsNout5
pOutvsNout6
pOutvsNout7
pOutvsNout8
pOutvsNout9
pOutvsNout10
pOutvsNout11
pOutvsNout12

```

```{r}
## Scale

z_scale <- function(column){
  column_c <- (column - mean(column)) / sd(column)
}

T1noOutS <- T1noOut %>% mutate_at(c("HR1", "HR2", "Resp1", "Resp2"), z_scale)
T2noOutS <- T2noOut %>% mutate_at(c("HR1", "HR2", "Resp1", "Resp2"), z_scale)
T3noOutS <- T3noOut %>% mutate_at(c("HR1", "HR2", "Resp1", "Resp2"), z_scale)

# Plot again to check how scaled data look like
pOutvsNout1S <- ggplot()+
  geom_line(data = T1noOutS, aes(time, HR1, color = "red"))+
  geom_line(data = T1noOut, aes(time, HR1, color = "blue"))

pOutvsNout2S <- ggplot()+
  geom_line(data = T1noOutS, aes(time, HR2, color = "red"))+
  geom_line(data = T1noOut, aes(time, HR2, color = "blue"))

pOutvsNout3S <- ggplot()+
  geom_line(data = T1noOutS, aes(time, Resp1, color = "red"))+
  geom_line(data = T1noOut, aes(time, Resp1, color = "blue"))

pOutvsNout4S <- ggplot()+
  geom_line(data = T1noOutS, aes(time, Resp2, color = "red"))+
  geom_line(data = T1noOut, aes(time, Resp2, color = "blue"))

pOutvsNout5S <- ggplot()+
  geom_line(data = T2noOutS, aes(time, HR1, color = "red"))+
  geom_line(data = T2noOut, aes(time, HR1, color = "blue"))

pOutvsNout6S <- ggplot()+
  geom_line(data = T2noOutS, aes(time, HR2, color = "red"))+
  geom_line(data = T2noOut, aes(time, HR2, color = "blue"))

pOutvsNout7S <- ggplot()+
  geom_line(data = T2noOutS, aes(time, Resp1, color = "red"))+
  geom_line(data = T2noOut, aes(time, Resp1, color = "blue"))

pOutvsNout8S <- ggplot()+
  geom_line(data = T2noOutS, aes(time, Resp2, color = "red"))+
  geom_line(data = T2noOut, aes(time, Resp2, color = "blue"))

pOutvsNout9S <- ggplot()+
  geom_line(data = T3noOutS, aes(time, HR1, color = "red"))+
  geom_line(data = T3noOut, aes(time, HR1, color = "blue"))

pOutvsNout10S <- ggplot()+
  geom_line(data = T3noOutS, aes(time, HR2, color = "red"))+
  geom_line(data = T3noOut, aes(time, HR2, color = "blue"))

pOutvsNout11S <- ggplot()+
  geom_line(data = T3noOutS, aes(time, Resp1, color = "red"))+
  geom_line(data = T3noOut, aes(time, Resp1, color = "blue"))

pOutvsNout12S <- ggplot()+
  geom_line(data = T3noOutS, aes(time, Resp2, color = "red"))+
  geom_line(data = T3noOut, aes(time, Resp2, color = "blue"))

pOutvsNout1S
pOutvsNout2S
pOutvsNout3S
pOutvsNout4S
pOutvsNout5S
pOutvsNout6S
pOutvsNout7S
pOutvsNout8S
pOutvsNout9S
pOutvsNout10S
pOutvsNout11S
pOutvsNout12S


```

```{R}
## Downsample
### This is tricky, so you can have a look at my code  (relying on Ludvig's groupdata2) if you get stuck
p_load(groupdata2)

df1 = T1noOutS %>%
  group(n = 100, method = 'greedy') %>%
  dplyr::summarise(
    time = mean(time,na.rm=T),
    HR1 = mean(HR1,na.rm=T),
    HR2 = mean(HR2,na.rm=T),
    Resp1 = mean(Resp1,na.rm=T),
    Resp2 = mean(Resp2,na.rm=T))

df2 = T2noOutS %>%
  group(n = 100, method = 'greedy') %>%
  dplyr::summarise(
    time = mean(time,na.rm=T),
    HR1 = mean(HR1,na.rm=T),
    HR2 = mean(HR2,na.rm=T),
    Resp1 = mean(Resp1,na.rm=T),
    Resp2 = mean(Resp2,na.rm=T))

df3 = T3noOutS %>%
  group(n = 100, method = 'greedy') %>%
  dplyr::summarise(
    time = mean(time,na.rm=T),
    HR1 = mean(HR1,na.rm=T),
    HR2 = mean(HR2,na.rm=T),
    Resp1 = mean(Resp1,na.rm=T),
    Resp2 = mean(Resp2,na.rm=T))

downsample <- function(data){
    data <- data %>% 
      group(n = 100, method = 'greedy') %>%
        dplyr::summarise(
        time = mean(time,na.rm=T),
        HR1 = mean(HR1,na.rm=T), 
        HR2 = mean(HR2,na.rm=T),
        Resp1 = mean(Resp1,na.rm=T),
        Resp2 = mean(Resp2,na.rm=T))
    return(data)
}

## Plot the downsampled data
p1 <- ggplot(data = df1) +
  geom_path(aes(time, Resp1, color = "P1")) +
  geom_path(aes(time, Resp2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")

p2 <- ggplot(data = df1) +
  geom_path(aes(time, HR1, color = "P1")) +
  geom_path(aes(time, HR2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")

p3 <- ggplot(data = df2) +
  geom_path(aes(time, Resp1, color = "P1")) +
  geom_path(aes(time, Resp2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")

p4 <- ggplot(data = df2) +
  geom_path(aes(time, HR1, color = "P1")) +
  geom_path(aes(time, HR2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")

p5 <- ggplot(data = df3) +
  geom_path(aes(time, Resp1, color = "P1")) +
  geom_path(aes(time, Resp2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")

p6 <- ggplot(data = df3) +
  geom_path(aes(time, HR1, color = "P1")) +
  geom_path(aes(time, HR2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")

p1
p2
p3
p4
p5
p6


## Now add the group, trial, condition to the cleaned up, scaled, downsampled data
## Tip the info is in the file name
df1$group <- "1"
df1$trial <- "1"
df1$condition <- "Synchronous"

df2$group <- "1"
df2$trial <- "2"
df2$condition <- "TurnTaking"

df3$group <- "1"
df3$trial <- "3"
df3$condition <- "SelfPaced"

addInfo <- function(df, filename){
  values <- str_extract_all(filename, "\\d")[[1]]
  df$group <- values[2]
  df$trial <- values[3]
  df$condition <- str_extract(filename, "(?<=_)[a-zA-Z]{2,}")
  return(df)
}
```


## Now we are ready to go to load and pre-process all files

Go through all the files (with a function passed onto map_df), check which files should be excluded, if any, and save the pre-processed time-series

A couple of tips:
- looping is oh so slow. Making a function and using Map/Map_df is your salvation.
- each study restarts the group numbering, so you should make sure to change that (e.g. 100 * Study + Group)
- you need to make sure all the data are meaningful or something has to be removed. Plotting is your friend. E.g. "Study1_G1_T1_Synchronous" has one bad respiration signal. We could replace it with NAs


```{r}
# Define a function running the loading, artifact removal, scaling, downsampling, info adding.
data_preprocess <- function(filename, inputdir, outputdir){
  print(filename)
  #create output folder, will only actually do something on the first function call
  inputfile <- file.path(inputdir, filename)
  dir.create("output", showWarnings = FALSE)
  outputfile = file.path(outputdir, filename)
  #check if files already in output
  if(file.exists(outputfile)){
    data <- read_csv(outputfile)
    return(data)
  }
  #Else
  data <- read_csv(inputfile)
  data <- data
  #check for study4
  if(str_detect(filename, "Study4")){
    colnames(data)[colnames(data)=="min"] <- "time"
  }
  data <- data %>% mutate_at(c("Resp1", "Resp2", "HR1", "HR2"), removeOuts)
  data <- data %>% mutate_at(c("Resp1", "Resp2", "HR1", "HR2"), z_scale)
  data <- downsample(data)
  data <- addInfo(data, filename)
  write_csv(data, outputfile)
  return(data)
}

#  Identify all files to be read
files <- list.files("data", pattern="\\.csv$")

# Run the function on the whole dataset using map_df
output <- purrr::map_df(files, data_preprocess, "data", "output")
        

# We realised, we had not extracted the study number. Since the function takes 13 hours to run again, we solve it in this way.

# Creating function, which groups numbers with consecutive numbers (in order to "recreate" the studies, which belong together in the output).
groupConsequtives <- function(al){
  #list of lists ;)
  lol <- list()
  listcoords <- which(diff(al) != 1)
  startCoord <- 1
  if(length(listcoords) == 0){
    return(list(al)) # Får tilbage som en liste
  }
  for(c in listcoords){
    newl <- list(al[startCoord:c]) # liste af alle værdier fra start til c (hver værdi i list-cords = index for når en gruppe slutter)
    startCoord <- c + 1
    lol <- c(lol, newl)
  }
  if(length(al) > listcoords[length(listcoords)])
  {
    lol <- c(lol, list(al[(listcoords[length(listcoords)]+1):length(al)]))
  }
  return(lol)
}

# Applying the function to the data
listofgroups <- groupConsequtives(output$.groups)

# Extracting study number from the filenames
study <- str_extract(files, "(\\d)")

# 
currentindex <- 0
output$Study <- 0
for (i in 1:length(listofgroups)){
  output[listofgroups[[i]] + currentindex,]$Study <- study[i]
  currentindex <- currentindex + length(listofgroups[[i]])
}

# FORDI VI HAR FUCKET OP I FUNKTIONEN - SÅ DET ER IKKE I DET ENDELIGE DOKUMENT (fordi 10 bliver taget som 2 tal og ikke som 1 tal...)
output[which(output$Study == 2 & output$trial == 0),]$group <- 10
output[which(output$Study == 3 & output$trial == 0),]$group <- 10

output[which(output$Study == 3 & output$group == 10 & output$condition == "Synchronous"),]$trial <- 1
output[which(output$Study == 3 & output$group == 10 & output$condition == "Conversation"),]$trial <- 2
output[which(output$Study == 3 & output$group == 10 & output$condition == "TurnTaking"),]$trial <- 3

output[which(output$Study == 2 & output$group == 10 & output$condition == "Conversation"),]$trial <- 1
output[which(output$Study == 2 & output$group == 10 & output$condition == "TurnTaking"),]$trial <- 2
output[which(output$Study == 2 & output$group == 10 & output$condition == "Synchronous"),]$trial <- 3

```

Now we need to make sure all the data are meaningful or something has to be removed
E.g. "Study1_G1_T1_Synchronous" has one bad respiration signal. We could replace it with NAs
```{r}
# Replacing NaNs with NAs
output[output == "NaN"] <- NA

# Unique ID per pair
output$Study <- as.numeric(output$Study)
output$pairID <- 100*output$Study+output$group

# Putting time on the same scale (seconds) (study 3 is different from the others) - assuming that the large values in study 3 are in milliseconds and that the rest are in minutes
output$time <- ifelse(output$time > 400, output$time/1000, output$time*60)

# Making a "time-since-the-beginning-of-the-experiment" column
output <- output %>%
  group_by(pairID, condition) %>%
  mutate(timesincestart = time - min(time))
```

Plotting to look for bad data aka identifying bad data
```{r}


  
ggplot(data = output[which(output$pairID == 102 & output$condition == "Synchronous"),]) +
  geom_line(aes(timesincestart, HR1, color = "HR1")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position = "bottom")

```


```{r}
# Remove bad data
testSeq <- function(ts, threshold){
  runningTot <- 0
  sequential <- rle(ts)
  for(i in 1:length(sequential$lengths)){
    if(sequential$lengths[i] > 10){
      interval <- output[sequential$lengths[i] + runningTot,]$timesincestart - output[1 + runningTot,]$timesincestart
      if(interval > threshold){
        ts[(1 + runningTot):(runningTot+sequential$lengths[i])] <- NA
      }
    }
    runningTot <- runningTot + sequential$lengths[i]
  }
  return(ts)
}

output <- output %>%
  mutate_at(c("HR1", "HR2", "Resp1", "Resp2"), testSeq, 2)

# Saving data
write_csv(output, "Preprocessed_data.csv")
```

## Now we need to run some analysis

Let's start with a multilevel model that accounts for 
- stability (how each signal is autocorrelated)
- interpersonal dependence (each signal is dependent from the previous state of the other signal)

The data needs to be further prepared, so we can analyze both participants in the same model.
We need to turn the data into a long format:
- a column indicating own hr and one own respiration
- a column indicating other hr and one other respiration
- a column indicating change in hr from previous round and one in respiration

We can then run an analysis where change is a function of one's previous state (stability, see slides), and the other's previous state (coupling). Make sure to:
- set up the most interesting contrasts: how do these parameters vary by condition? which condition should be baseline?
- set up the right random effects.
- N.B. the model will be slow. Make sure it works on a subset of the data first!

Bonus question: what if we include an additional layer? Is my heart rate just adjusting to yours, or also to how much you are adjusting to mine?
- to start answering this we can add a column indicating the previous change in hr in the other and one in respiration
- we can then build on the previous models by also adding the previous change in the other


```{r}
# Genearate a column for each: previous HR1, HR2, Resp1, Resp2
output$prevHR1 <- lag(output$HR1, 1)
output$prevHR2 <- lag(output$HR2, 1)
output$prevResp1 <- lag(output$Resp1, 1)
output$prevResp2 <- lag(output$Resp2, 1)

# Genearate a column for each: change in HR1, HR2, Resp1, Resp2
output$changeHR1 <- output$HR1 - output$prevHR1
output$changeHR2 <- output$HR2 - output$prevHR2
output$changeResp1 <- output$Resp1 - output$prevResp1
output$changeResp2 <- output$Resp2 - output$prevResp2

# Make the data long, so we can analyze both participants at the same time 
## N.B. This is a bit tricky and you might have to do it in several steps
output %>% 
  gather(
    participant, change_HR_self,
    changeHR1, changeHR2)
  select()
  )

d_hr_change
gather (d, # data
participant, HR_change_self , # new vars
HR1_change, HR2_change) %>% #old vars
select ( # drop irrelevant vars
time, HR_change_self , participant,
Study , Group, Condition , Trial)
mutate ( # create unique participant ID
participant = parse_number as.character (Group)) * 10 +
parse_number (


# Set the most interesting contrast e.g. by defining synchronous or conversation as the baseline

# Model change as a function of own and other previous state 


# Bonus points: Add to the previous model also change in the other to see whether my adaptation is influenced by the other's adaptation.

```


## Now we need to create control baselines.

First shuffled controls, then surrogate pairs.

### Creating controls: shuffled controls

Shuffled controls break the temporal dependencies of time-series by shuffling the value within one time-series. This ensures the "coordination" observed is not due to the actual values in the series and not their sequence.
Tip: sample() is your friend, but make sure to shuffle things within participant/condition and not throughout the whole dataset
 
```{r}

# Create a shuffled dataset
# Concatenate it to the original dataset (and remember to have a column telling you which is which)

# Create the same models as in the previous chunk, but adding an interaction by shuffled vs. real

```
 
 
### TRICKY! Creating controls: surrogate pair controls
 - Per each real pair, identify at least one surrogate pair (matching one of the participants, with somebody doing the same task, but in a different pair)

```{r}

# Identify unique pairs within a given study (to keep things manageable) and create list of possible surrogate pairs (e.g. individual 1 from pair 1 and individual 2 from pair 2)

# Starting from the wide format, create "surrogate" dataset with the data from surrogate pairs

# Make it into long format

# Create models as in chunks above, but adding an interaction with the Real vs. Surrogate variable (exclude shuffled ones for simplicity)



```
 

### Effects of respiration coordination on heart rate coordination
 - describe how you would test those.
 - Optional: run the models and report them

 